---
title: "Mini project by Cedric"
author: "Cedric Xavier M"
date: "`r Sys.Date()`"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = TRUE)
```

```{r}
getwd()
```

# MINI PROJECT PRESENTATION

The Mini project on the Maternal Health RisK Dataset
### Load the Maternal health data set 
# load the various packages that will be used

```{r}

library(dplyr)
library(tidyverse)
library(dslabs)
```


```{r}
library(reticulate)

```
```{python}
import sys
print(sys.version)
```
```{r}
library(reticulate)
np <- import("numpy")
plt <- import("matplotlib.pyplot")
pd<-import("pandas")
Maternal_health<- read.csv( file= 'Maternal Health Risk Data Set.csv' )
```

## print the data set
```{r}
Maternal_health
```


## Grouping by Risk level and finding the mean age of each risk group in descending order 

```{r}
Maternal_health %>% 
  group_by(RiskLevel) %>%
  summarize(mean_age= mean(Age))%>% 
  arrange(desc(mean_age))
```
- We observe an increase of the risk level with age

## Median Diastolic BP by risk group


```{r}
Maternal_health %>% 
  group_by(RiskLevel) %>%
  summarize(median_Diastolic= median(DiastolicBP))%>% 
  arrange(median_Diastolic)
```
- High group has a higher diastolic BP 
- both analysis suggest that older individuals are at a higher risk compared to younger individuals also higher Diastolic BP is correlated to high risk
-This may suggest that diastolic pressure is higher in older individuals and drives level of risk

## Creating  a new variable **systolic rate**. 
-systolic rate = systolicBP/age

```{r}
Maternal_health1 <- mutate(Maternal_health, systolic_rate = SystolicBP )
head(Maternal_health1)
```
##  Filter participants of age <= 40 years and determine average systolic rate for each risk level

```{r , echo = TRUE}
Filtered_by_age <- filter(Maternal_health1, Age <= 40)
print(Filtered_by_age)
```

## Determine average systolic rate for each risk level
```{r}
Filtered_by_age %>% 
  group_by(RiskLevel) %>%
  summarize(average_systolic_rate= mean(systolic_rate)) 
```
- Individuals at higher risk level have a higher systolic rate
- Just like Diastolic BP systolic rate increases with risk level

## calculate the five-number summary for each variable
```{r}
py$Maternal_health <- Maternal_health1
```


```{python}
import numpy as np
import pandas as pd
print(r.Maternal_health1)
df= r.Maternal_health1
```
## Create histogram and box plot for all continious variable and calculate the five-number summary for each variable, grouped by risk level

## Histograms for all continious variable

### Histogram 

```{python}
import seaborn as sns
import matplotlib.pyplot as plt
import numpy as np
import pandas as pd
df= r.Maternal_health
continous_variables = ['Age', 'SystolicBP', 'DiastolicBP', 'BS', 'BodyTemp', 'HeartRate']
for var in continous_variables:
  plt.figure()
  sns.histplot(df[var], kde=True) 
  plt.title(f'Histogram of {var}')
  plt.xlabel(var)
  plt.ylabel('Frequency')
  plt.show()



```
## Inside into the distribution of the data

- Is skewed to the left
- The 15 to 35 age group is the most represented
- Diastolic BP, systolic BP and Heart rate are normally distributed


## Box plots
```{python}
import seaborn as sns
import matplotlib.pyplot as plt
import numpy as np
import pandas as pd
df= r.Maternal_health
continous_variables = ['Age', 'SystolicBP', 'DiastolicBP', 'BS', 'BodyTemp', 'HeartRate']
for var in continous_variables:
  plt.figure()
  sns.boxplot(data=df[continous_variables])  
  plt.title('Box Plots of Continous Variables')
  plt.show()
  

```
plt.boxplot(data=df[var] , notch =False , vert = True , widths = 0.2 , labels = df[var]
            ,showmeans = True , boxprops = dict(color= "red") ,capprops = dict(color = "red") , sym= "g+",  whiskerprops = dict(color = "red") )

  plt.show()

## calculate the five-number summary for each variable, grouped by risk level]

```{python}
import seaborn as sns
import matplotlib.pyplot as plt
import numpy as np
import pandas as pd
df= r.Maternal_health
byRiskLevel = df.groupby('RiskLevel')
```

###  age by riskLevel 
```{python}
import numpy as np
import pandas as pd
byRiskLevel = df.groupby('RiskLevel')
byRiskLevel.describe()['Age']
```

### -Systolic BP by RiskLevel
```{python}
import numpy as np
import pandas as pd
byRiskLevel = df.groupby('RiskLevel')
byRiskLevel.describe()['SystolicBP']


```
###Diastolic BP by RiskLevel

```{python}
import numpy as np
import pandas as pd
byRiskLevel = df.groupby('RiskLevel')
byRiskLevel.describe()['DiastolicBP']
```
### BS RiskLevel

```{python}
import numpy as np
import pandas as pd
byRiskLevel = df.groupby('RiskLevel')
byRiskLevel.describe()['BS']

```
### Body Temperature by Risk level
```{python}
import numpy as np
import pandas as pd
byRiskLevel = df.groupby('RiskLevel')
byRiskLevel.describe()['BodyTemp']

```
### Heart rate  by risk level
```{python}
import numpy as np
import pandas as pd
byRiskLevel = df.groupby('RiskLevel')
byRiskLevel.describe()['HeartRate']

```

## Realtionship between variables

### Age as an independent variable

```{r}
library(dslabs)
Maternal_health
plot(Maternal_health$Age, Maternal_health$HeartRate )
plot(Maternal_health$SystolicBP, Maternal_health$HeartRate )
plot(Maternal_health$DiastolicBP, Maternal_health$HeartRate )
plot(Maternal_health$BodyTemp, Maternal_health$HeartRate )
```
- There seem to be no correlation between age does not influence Systolic nor Diastolic BPs, nor body temperature



```
- There is no relationship between body temperature,systolic and diastolic BP


```{r}

```


